Enabling Ambient Backscatter \\Using Low-Cost Software Defined Radio

### 2017-04-07 paper.tex
Worked on the transmitter part.

\subsection{Transmitter}
\begin{figure}[h]
\centering
\includegraphics[width=0.6\columnwidth]{./fig/s11v2}
\caption{Input reflection coefficient \ensuremath{\Gamma_{in}} also known as \ensuremath{S_{\text{11}}} of the self-made backscattering ground-plane antenna.}
\label{fig:s11}
\end{figure}
A backscattering antenna is the most important part of backscatter transmitter. Our antenna is optimized for usage with the television signal, which has its center frequency at 626 MHz. It is a ground plane antenna, which has been built up on a large unetched PCB plate. This rectangular PCB, with dimensions of 200 mm x 300 mm, is covered with a copper layer on one side and left blank on the other. A hole in the middle of the PCB serves for connecting a wire to a 50 \ensuremath{\Omega} micro strip transmission line. The wire, which acts as the vertical antenna element and has a length of 350 mm (roughly 0.73 \ensuremath{\lambda}), is soldered to the transmission line. The transmission line has been calculated for 626 MHz using an FR4 substrate with a thickness of 1.6 mm. According to micro strip line theory the length of the line changes the phase and the width of the line changes the impedance (cf. Chapter 3 in \cite{pozar_microwave_2011}). The design equations given in \cite{pozar_microwave_2011} are implemented in various programs e.g. KiCAD, which has been used for solving the design equations. The given parameters lead to a micro strip line width of 3 mm. An open-stub, attached directly at the juncture where the vertical antenna element is attached to the transmission line, is used for fine tuning of the input impedance. The stub length has been adjusted manually (shortened) while measuring the reflection coefficient with a vector network analyzer (VNA). Open-stub matching is discussed in Chapter 5 in \cite{pozar_microwave_2011}. Furthermore the resonance frequency has been tuned by cutting the vertical antenna element piece by piece while measuring with the VNA until \ensuremath{S_{\text{11}}} had the minimum close to the desired frequency (Figure~\ref{fig:s11}\cp{Can we replace all these figures by PDFs?}). In the plot one can also see, that the antenna is quite narrowband\cp{We should report something like the 3dB bandwidth rather than saying something non-specific like this}. For backscattering this is not necessarily a disadvantage, since everything that is within the operating frequency of the antenna is backscattered. With our narrowband antenna the undesired backscattering of out-of-band signals is reduced. With our approach we were able to achieve an input impedance at 634 MHz of \ensuremath{Z_{in}=(52.2 + j 4.3) \Omega}. At 626 MHz \ensuremath{S_{\text{11}}} is still approximately -17 dB.

Besides the antenna, the transmitter consists of an \textit{MSP430} MCU, that sends data to the receiver. This transmitter controls an RF switch through an I/O pin. The switch can alternate the impedance connected to the antenna between matched (50 \ensuremath{\Omega}) and open. The software now steers this switch with a two 2 MHz rectangular timer signal to transmit a data 1 and is just leaves it open to transmit a data 0. Therefore the result is a frequency shift\cp{We need to explain this in the background} of the television signal by 2 MHz for a 1 and no shift for a 0. Hence a classical binary amplitude shift keying is the implemented modulation technique.

>>>>

\subsection{Transmitter}
\begin{figure}[h]
\centering
\includegraphics[width=0.6\columnwidth]{./fig/s11v2}
\caption{Input reflection coefficient \ensuremath{\Gamma_{in}} also known as \ensuremath{S_{\text{11}}} of the self-made backscattering ground-plane antenna.}
\label{fig:s11}
\end{figure}

A backscattering antenna is the most important part of backscatter
transmitter. Our antenna is optimized for usage with the television
signal, which has its center frequency at 626 MHz. It is a ground plane
antenna, which has been built up on a large unetched PCB plate. A hole
in the middle of the PCB serves for connecting a wire to a 50
\ensuremath{\Omega} micro strip transmission line. The wire, which acts
as a vertical antenna element is soldered to the transmission line. The
transmission line has been calculated for 626 MHz using an FR4 substrate
with a thickness of 1.6 mm. The design equations given in
\cite{pozar_microwave_2011} are implemented in various programs e.g.
KiCAD, which has been used for solving the design equations. An
open-stub, attached directly at the juncture where the vertical antenna
element is attached to the transmission line, is used for fine tuning of
the input impedance (cf. chapter 5 in \cite{pozar_microwave_2011}).
Accepting -15 dB as sufficient value for \ensuremath{S_{11}}, one can
say, that the antenna works in a range from 626 MHz to 645 MHz. For
backscattering this small bandwidth is not necessarily a disadvantage,
since everything that is within the operating frequency of the antenna
is backscattered. With our narrowband antenna the undesired
backscattering of out-of-band signals is reduced.  With our approach we
were able to achieve an input impedance at 634 MHz of
\ensuremath{Z_{in}=(52.2 + j 4.3) \Omega}. At 626 MHz
\ensuremath{S_{\text{11}}} is still approximately -17 dB.

Besides the antenna, the transmitter consists of an \textit{MSP430} MCU,
that sends data to the receiver. This transmitter controls an RF switch
through an I/O pin. The switch can alternate the impedance connected to
the antenna between matched (50 \ensuremath{\Omega}) and open. The
software now steers this switch with a two 2 MHz rectangular timer
signal to transmit a data 1 or it just leaves the antenna open to
transmit a data 0. Therefore the result is a frequency shift of the
television signal by 2 MHz for a 1 and no shift for a 0. This can be
easily understood thinking about a simple cosine, which is shifted by
another cosine: \ensuremath{2\,cos(a)cos(b)=cos(a-b)+cos(a+b)}. Of
course this example is simplified compared to the actual situation as
the actual signal, which is used for shifting contains more than one
frequency as well as the TV signal, which is shifted. Hence a classical
binary amplitude shift keying is the implemented modulation technique.
